### 2010-03-24 papers/icfp-2010-desc/paper_idesc.tex
Desc@ICFP: fix GADT discussion.

The $\Fin$ example is also a well-known example of Generalised
Abstract Data-Type (GADT) in Haskell. It comes as no surprise that our
universe of indexed description supports this restricted class of
dependent data-types. As we have shown with $\Fin$, constrained
constructors are straightforwardly translated into equality
constraints.

While we support the same grammar, indexed descriptions embraces a
much wider class of data-types. Indeed, we pose no constraint apart
from strict positivity\note{pwm: Is this actually the 1st mention of this
phrase? I think that should be rectified :)}. This suggests that, based on
the familiar grammar of GADTs, we can simply lift off the restrictions and
translates the user's input to a code in our universe of indexed description.

However, one could, quite rightfully, be worried by the introduction
of equations in the picture. In the context of programming, these
equations will come back and haunt us, begging for proofs. Hence,
whereas the user typed a constructor-constrained data-type, the
elimination form will involve equations and their proofs. This
introduces a loophole in our abstraction we should try to address.

\note{Is that convincing? Worth saying? Anything else?}

@@

To get a clue of a solution, let us look back at $\Fin$. We note that
the equation is introduced because we are \emph{storing} the index of
the inductive family, through $m$. Hence the constraint. However,
\emph{inductive families need not store their
  indices}~\cite{brady:index-inductive-families}\note{ I'm not
  familiar with this, I need to get into the paper. }. Applying the
\emph{forcing} optimisation to our definition of $\Fin$ gives the
following, equivalent definition:

@@

Equations have simply disappeared! We should precise that forcing a
description is not guaranteed to remove all constraints. It is subject
of future work to see if constraints can be entirely eradicated, or
presented more conveniently to the user. Finally, it is worth
mentioning that the forcing optimisation is a \emph{source-to-source}
transformation of the description.

>>>>

The $\Fin$ example is also a well-known example of Generalised
Abstract Data-Type (GADT) in Haskell. It comes as no surprise that our
universe of indexed description supports this restricted class of
inductive families. As we have seen with $\Fin$, constrained
constructors are straightforwardly translated into equations.

However, one could, quite rightfully, be worried by the introduction
of equalities in the picture. In a programming context, equalities
haunt us, begging for proofs. Hence, whereas the user typed a
constructor-constrained data-type, the elimination form will involve
equalities and their proofs. As much as possible, one would rather
avoid suffering from this burden. However, the equalities introduced
here are rather special: they are \emph{equations} constraining
variables to certain terms. Goguen et
al.~\cite{goguen:pattern-matching} have shown, in the context of
dependent pattern matching, how to tame these equations.

@@

Further, we can actually remove these equations altogether. Let us
look back at $\Fin$. We note that the equations are introduced because
we are \emph{storing} the index of the inductive family, through
$m$. Hence the constraints. However, \emph{inductive families need not
  store their indices}~\cite{brady:index-inductive-families}. Applying
the \emph{forcing} optimisation to our initial definition of $\Fin$
gives the following, equivalent definition:

@@

Equations have simply disappeared. We should precise that forcing a
description is not guaranteed to remove all constraints. It is subject
to future work to see if constraints can be entirely eradicated, or
presented more conveniently to the user. Finally, it is worth
mentioning that the forcing optimisation is a \emph{source-to-source}
transformation of the description.
